Sparse Function-space Representation \\ of Neural Networks

### 2023-05-17 paper/main.tex
commetned out because we don't show this

\paragraph{\our for dual conditioning} Adding new data to a trained neural network without retraining from scratch is a non-trivial problem. Some approaches exist \citep{kirsch2022marginal, spiegelhalter1990sequential} but are in the weight space formulation, which makes conditioning harder to formulate. As pointed out in \cite{chang2022fantasizing}, dual parameter conditioning updates are straightforward and were the original motivation of \cite{csato2002sparse}. This involves integrating data into the sum in \cref{eq:dual_sparse}. Thus, we need to find the data dual variables $\hat{\alpha}$ and $\hat{\beta}$ for the new data subset $\mathcal{D}_\textrm{new} = \{(\vx_i,\vy_i)\}_{i=1}^{N_{\textrm{new}}}$ and compute the corresponding kernel terms. As pointed out in \cite{chang2022fantasizing}, for GPs, these updates are exact under a Gaussian likelihood and the kernel remains constant. For non-Gaussian likelihoods, the results are almost the same as retraining with all the data. In practice, given our NN-derived kernel's dependency on the Jacobian of the trained neural network $\Jac{\weights_*}{\vx}$, the Jacobians become outdated as we add more data. However, we can still improve predictions by conditioning data without retraining. 
%[SEE EXPERIMENT ???].
%%The resulting regularizer can be computed as follows
%%\begin{equation}
%%	\mathcal{R}_\textit{SFR}(\weights, \mathcal{M}) = \sum_{s=1}^{t-1} \frac{1}{m} \left[\left(f_{\weights}(\MZ_{s}) - f_{\weights_{s}}(\MZ_s) \right)^\T \bar{\MB}^{-1}_{s} \left(f_{\weights}(\MZ_{s}) - f_{\weights_{s}}(\MZ_s) \right) \right].
%%\end{equation}
%%
%\begin{equation}
%	\mathcal{R}_\mathrm{\our}(\weights, \mathcal{M}) = \frac{1}{2} \sum_{s=1}^{t-1} \frac{1}{M} 
%	\left\lVert 
%	f_{\weights}(\MZ_{s}) - \vu_t % f_{\weights_{s}}(\MZ_s)
%	\right\rVert_{\bar{\MB}^{-1}_{s}}.
%\end{equation}
%%\todo{choose one option}
%% where $f_\vw(\MZ_{s}) \in \R^m$ are the outputs of neural networks with the current weights $\vw$ and stored function outputs of the neural network at the MAP optimum for task $s$, i.e., each $i$th entry is $f_\cdot(\vz_i) \in \R$.
%%\begin{equation}
%%	\argmin_{\weights} \sum_{i=1}^{n} \ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i) + \delta \mathcal{R}(\weights) + \tau \mathcal{R}_\textit{SFR}(\weights)
%%\end{equation}
%Then, when the training on the successive task, the new objective to minimize takes the following form
%\begin{equation}
%	\weights_t^* = \argmin_{\weights} \mathcal{L}(\mathcal{D}_t, \weights) + \tau \mathcal{R}_\textit{SFR}(\weights, \mathcal{M})
%\end{equation}

\section{Experiments}
\label{sec:experiments}
%
We present a series of experiments specifically designed to showcase the robustness and effectiveness of \our. The experiments are aimed at answering the following key questions:
\textbf{Predictions:}~Does our method's ability to consider the full data set offer benefits over subset function-space methods? (\cref{sec:uci}) How do predictions with our sparse function-space approximation compare to weight-space approximations? 
\textbf{Function-space updates:} How fast are our function-space updates relative to retraining from scratch? Do they improve predictive performance? 
\textbf{Representation:} Can a sparse function-space representation help retain knowledge from previous tasks in CL? (\cref{sec:cl-exp})
\textbf{Uncertainty:} Can our uncertainty estimates help exploration in RL? (\cref{sec:rl-exp}) We implement all methods in PyTorch~\cite{paszke2019pytorch} and run the benchmarks using a GPU cluster. We provide full experiment details in \cref{app:experiments}.

@@

\begin{table}[t!] 
  \centering\scriptsize
  \caption{Metrics for supervised learning on image data with a CNN. We report accuracy, NLPD, and expected calibration error (mean$\pm$std) over 10 seeds. Our \our method is able to match the full methods using only a sparse set. For the GP predictive sparse subset method and \our we report results both using the NN MAP as mean (NN) and GP  mean.} 
	\label{tbl:imagesuper}
	% Control table spacing
	\renewcommand{\arraystretch}{1.}
	\setlength{\tabcolsep}{6pt}
	\setlength{\tblw}{0.095\textwidth}

@@

    GP predictive NN & $1024$ & \val{97.70}{0.07} & \val{0.080}{0.002} & \val{0.011}{0.001}
                              & \val{91.47}{0.41} & \val{0.289}{0.009} & \val{0.033}{0.003} \\

    GP predictive & $1024$  & \val{76.28}{1.85} & \val{0.702}{0.029} & \val{0.070}{0.022}
                            & \val{83.98}{0.35} & \val{0.503}{0.012} & \val{0.052}{0.006} \\
    \our NN & $1024$ &  \val{98.02}{0.12} & \val{0.066}{0.003} & \val{0.007}{0.001}
                     & \val{91.93}{0.45} & \val{0.267}{0.011} & \val{0.028}{0.003} \\
    \our & $1024$ & \val{98.07}{0.06} & \val{0.063}{0.003} & \val{0.005}{0.001}
                  & \val{91.56}{0.35} & \val{0.253}{0.006} & \val{0.012}{0.001} \\
    \midrule
%    \multirow{2}{*}{FMNIST}
%    & MAP & \val{91.39}{0.11} & \val{0.258}{0.004} & \val{0.017}{0.001} \\
    %& BNN predictive \cite{immer2021improving} & \val{84.42}{0.12} & \val{0.942}{0.016} & \val{0.411}{0.008} \\
    % & BNN predictive \cite{ritter2018kfac} & \val{91.20}{0.07} & \val{0.265}{0.004} & \val{0.024}{0.002} \\
    % & GLM predictive \cite{immer2021improving} & \val{92.25}{0.10} & \val{0.244}{0.003} & \val{0.012}{0.003} \\
    %& GP predictive \cite{immer2021improving} $m=3200$ & \val{91.36}{0.11} & \val{0.25}{0.004} & \val{0.007}{0.001} \\
    % & GP predictive NN $m=1024$ &  \val{0}{0.07} & \val{0}{0.002} & \val{0}{0.001} \\
    % & GP predictive $m=1024$  &  \val{0}{1.85} & \val{0}{0.029} & \val{0}{0.022} \\
    % & \our NN $m=1024$ &  \val{0}{0.0} & \val{0}{0.003} & \val{0}{0.001} \\
    % & \our $m=1024$ &  \val{0}{0.06} & \val{0}{0.003} & \val{0}{0.001} \\
    \bottomrule
  \end{tabular}

>>>>

\paragraph{\our for dual conditioning} Adding new data to a trained neural network without retraining from scratch is a non-trivial problem. Some approaches exist \citep{kirsch2022marginal, spiegelhalter1990sequential} but are in the weight space formulation, which makes conditioning harder to formulate. As pointed out in \cite{chang2022fantasizing}, dual parameter conditioning updates are straightforward and were the original motivation of \cite{csato2002sparse}. This involves integrating data into the sum in \cref{eq:dual_sparse}. Thus, we need to find the data dual variables $\hat{\alpha}$ and $\hat{\beta}$ for the new data subset $\mathcal{D}_\textrm{new} = \{(\vx_i,\vy_i)\}_{i=1}^{N_{\textrm{new}}}$ and compute the corresponding kernel terms. As pointed out in \cite{chang2022fantasizing}, for GPs, these updates are exact under a Gaussian likelihood and the kernel remains constant. For non-Gaussian likelihoods, the results are almost the same as retraining with all the data. However, it is worth noting that our kernel's dependency on the trained neural network $\Jac{\weights_*}{\vx}$ (via the Jacobian) becomes outdated as we add more data. Nevertheless, we can still improve predictions by conditioning on new data without retraining.
%[SEE EXPERIMENT ???].
%%The resulting regularizer can be computed as follows
%%\begin{equation}
%%	\mathcal{R}_\textit{SFR}(\weights, \mathcal{M}) = \sum_{s=1}^{t-1} \frac{1}{m} \left[\left(f_{\weights}(\MZ_{s}) - f_{\weights_{s}}(\MZ_s) \right)^\T \bar{\MB}^{-1}_{s} \left(f_{\weights}(\MZ_{s}) - f_{\weights_{s}}(\MZ_s) \right) \right].
%%\end{equation}
%%
%\begin{equation}
%	\mathcal{R}_\mathrm{\our}(\weights, \mathcal{M}) = \frac{1}{2} \sum_{s=1}^{t-1} \frac{1}{M} 
%	\left\lVert 
%	f_{\weights}(\MZ_{s}) - \vu_t % f_{\weights_{s}}(\MZ_s)
%	\right\rVert_{\bar{\MB}^{-1}_{s}}.
%\end{equation}
%%\todo{choose one option}
%% where $f_\vw(\MZ_{s}) \in \R^m$ are the outputs of neural networks with the current weights $\vw$ and stored function outputs of the neural network at the MAP optimum for task $s$, i.e., each $i$th entry is $f_\cdot(\vz_i) \in \R$.
%%\begin{equation}
%%	\argmin_{\weights} \sum_{i=1}^{n} \ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i) + \delta \mathcal{R}(\weights) + \tau \mathcal{R}_\textit{SFR}(\weights)
%%\end{equation}
%Then, when the training on the successive task, the new objective to minimize takes the following form
%\begin{equation}
%	\weights_t^* = \argmin_{\weights} \mathcal{L}(\mathcal{D}_t, \weights) + \tau \mathcal{R}_\textit{SFR}(\weights, \mathcal{M})
%\end{equation}

\section{Experiments}
\label{sec:experiments}
%
We present a series of experiments specifically designed to showcase the robustness and effectiveness of \our. The experiments are aimed at answering the following key questions:
\textbf{Predictions:}~Does our method's ability to consider the full data set offer benefits over subset function-space methods? (\cref{sec:uci}) How do predictions with our sparse function-space approximation compare to weight-space approximations? 
% \textbf{Function-space updates:} How fast are our function-space updates relative to retraining from scratch? Do they improve predictive performance?
\textbf{Representation:} Can a sparse function-space representation help retain knowledge from previous tasks in CL? (\cref{sec:cl-exp})
\textbf{Uncertainty:} Can our uncertainty estimates help exploration in RL? (\cref{sec:rl-exp}) We implement all methods in PyTorch~\cite{paszke2019pytorch} and run the benchmarks using a GPU cluster. We provide full experiment details in \cref{app:experiments}.

@@

\begin{table}[t!] 
  \centering\scriptsize
  \caption{Metrics for supervised learning on image data with a CNN. We report accuracy, NLPD, and expected calibration error (mean$\pm$std) over 10 seeds. Our \our method is able to match the full methods using only a sparse set. For the GP predictive sparse subset method and \our we report results both using the NN MAP as mean (NN) and using the GP mean.} 
	\label{tbl:imagesuper}
	% Control table spacing
	\renewcommand{\arraystretch}{1.}
	\setlength{\tabcolsep}{6pt}
	\setlength{\tblw}{0.095\textwidth}

@@

    GP predictive (NN) & $1024$ & \val{97.70}{0.07} & \val{0.080}{0.002} & \val{0.011}{0.001}
                              & \val{91.47}{0.41} & \val{0.289}{0.009} & \val{0.033}{0.003} \\

    GP predictive & $1024$  & \val{76.28}{1.85} & \val{0.702}{0.029} & \val{0.070}{0.022}
                            & \val{83.98}{0.35} & \val{0.503}{0.012} & \val{0.052}{0.006} \\
    \our (NN) & $1024$ &  \val{98.02}{0.12} & \val{0.066}{0.003} & \val{0.007}{0.001}
                     & \val{91.93}{0.45} & \val{0.267}{0.011} & \val{0.028}{0.003} \\
    \our & $1024$ & \val{98.07}{0.06} & \val{0.063}{0.003} & \val{0.005}{0.001}
                  & \val{91.56}{0.35} & \val{0.253}{0.006} & \val{0.012}{0.001} \\
%    \multirow{2}{*}{FMNIST}
%    & MAP & \val{91.39}{0.11} & \val{0.258}{0.004} & \val{0.017}{0.001} \\
    %& BNN predictive \cite{immer2021improving} & \val{84.42}{0.12} & \val{0.942}{0.016} & \val{0.411}{0.008} \\
    % & BNN predictive \cite{ritter2018kfac} & \val{91.20}{0.07} & \val{0.265}{0.004} & \val{0.024}{0.002} \\
    % & GLM predictive \cite{immer2021improving} & \val{92.25}{0.10} & \val{0.244}{0.003} & \val{0.012}{0.003} \\
    %& GP predictive \cite{immer2021improving} $m=3200$ & \val{91.36}{0.11} & \val{0.25}{0.004} & \val{0.007}{0.001} \\
    % & GP predictive NN $m=1024$ &  \val{0}{0.07} & \val{0}{0.002} & \val{0}{0.001} \\
    % & GP predictive $m=1024$  &  \val{0}{1.85} & \val{0}{0.029} & \val{0}{0.022} \\
    % & \our NN $m=1024$ &  \val{0}{0.0} & \val{0}{0.003} & \val{0}{0.001} \\
    % & \our $m=1024$ &  \val{0}{0.06} & \val{0}{0.003} & \val{0}{0.001} \\
    \bottomrule
  \end{tabular}
